---
title: "Ghana prediction"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      warning=FALSE, message=FALSE)
knitr::opts_knit$set(root.dir="../")
```

```{r load_data, echo=FALSE}
library(MLlibrary)
library(dplyr)

NAME <- "ghana"
mx <- load_dataset(NAME)
```

## Data

This is a sample analysis of the ghana dataset from 2006. We are just using the `r ncol(gh)` variables which went into a PMT which was recently developed by the WB. The dataset has `r nrow(gh)` households which have no missing data.

Least squares can be considered the benchmark method upon which the PMT from the WB was based.

Our target variable is log of (price adjusted) per (equivalence scale) capita consumption.

```{r load_models, echo=FALSE}
all_models <- load_models("mexico2")
```


## Simple Regression

```{r select_regression}
models <- c("least_squares", "ridge", "lasso", "rtree", "stepwise", "forest")
simple_regression <- filter(all_models, method %in% models)
```

```{r plot_simple_regression, echo=FALSE}
plot_swf_(simple_regression)
plot_reach_vs_waste_(simple_regression)
plot_reach_vs_waste_(simple_regression, THRESHOLD=.2)
```

```{r calculate_mse, echo=TRUE}
calculate_mse_(simple_regression)
```

## Simple Classification
```{r select_classifications}
models <- c("logistic_20", "ctree_20", "cforest_20", "logistic_30", "ctree_30", "cforest_30")
simple_classification <- filter(all_models, method %in% models | method == "least_squares")
```

```{r plot_simple_classification, echo=FALSE}
plot_swf_(simple_classification)
plot_reach_vs_waste_(simple_classification)
plot_reach_vs_waste_(simple_classification, THRESHOLD=.2)
plot_reach_vs_waste_(simple_classification, THRESHOLD=.3)
```

## Compound Methods

```{r select_compound}
# models <- c("ridge_rural" "mca_knn", "pca_knn", "mca_pca_avg")
# compound <- filter(all_models, method %in% models | method == "least_squares")
# ```
# 
# ```{r plot_compound, echo=FALSE}
# plot_swf_(compound)
# plot_reach_vs_waste_(compound)
# plot_reach_vs_waste_(compound, THRESHOLD=.2)
```
