```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      warning=FALSE, message=FALSE)
```

```{r load_data, echo=FALSE}
source("tanzania.R")
require(doMC)
registerDoMC(cores=3)

year <- 2008
tz08 <- create_dataset(2008)
```

## Data

This is a sample analysis of the tanzania dataset from `r year`.
The dataset has `r nrow(tz08)` rows and `r ncol(tz08)` columns.
Our goal is to predict the consumption levels from survey data.
In this case we take the log of per capita consumption as our target variable.

## Models

We trained a number of different models on the raw data.
We use `r k=5` fold cross validation for testing predictive accuracy.
```{r create_models}
tz08 <- standardize_predictors(tz08, "lconsPC")
x <- model.matrix(lconsPC ~ .,  tz08)
y <- tz08[rownames(x), "lconsPC"]

ridge <- kfold(k, Ridge(), y, x)
least_squares <- kfold(k, LeastSquares(), y, x)
```

We also trained each model on a dataset including all pairwise interactions.

```{r threshold, echo=FALSE}
threshold <- 12.5
percentile <- sum(tz08$lconsPC < threshold) / nrow(tz08)
```

```{r interactions}
# x_ix <- model.matrix(lconsPC ~ . + .:.,  tz08)
# y_ix <- tz08[rownames(x_ix), "lconsPC"]
# 
# least_squares_ix <- kfold(k, LeastSquares(), y_ix, x_ix)
```

## Assessments

### Density
```{r density, echo=FALSE}
plot_density(
  ridge=ridge,
  least_squares=least_squares)
```
Note that the following metrics take a threshold parameter to determine the true poor.
Here set the threshold to be `r threshold`, about the `r sprintf("%.2f", percentile)` percentile.

### ROC

```{r roc, echo=FALSE}
plot_roc(
  threshold,
  ridge=ridge,
  least_squares=least_squares)
```

### Accuracy (Coverage rate)
```{r accuracy, echo=FALSE}
plot_accuracy(
  threshold,
  ridge=ridge,
  least_squares=least_squares,
  SHOW_TRUE=TRUE,
  POINT_COUNT=50)
```

We can also display coverage rates at the consumption threshold and at the desired percentile.

```{r accuracy_cutoffs, echo=FALSE}
plot_accuracy(
  threshold,
  ridge=ridge,
  least_squares=least_squares,
  SHOW_CUTOFFS=TRUE,
  POINT_COUNT=50)
```

### Budget to the true poor
```{r accuracy_dollars, echo=FALSE}
plot_accuracy_dollars(
  threshold,
  ridge=ridge,
  least_squares=least_squares,
  POINT_COUNT=50)
```

### Social welfare
```{r swf, echo=FALSE}
plot_swf(
  ridge=ridge,
  least_squares=least_squares,
  GAMMA=2,
  POINT_COUNT=50)
```

### Leakage
```{r leakage, echo=FALSE}
plot_metric(
  leakage,
  ridge=ridge,
  least_squares=least_squares)
```

### Undercoverage
```{r undercoverage, echo=FALSE}
plot_metric(
  undercoverage,
  ridge=ridge,
  least_squares=least_squares)
```

### nCGH
```{r ncgh, echo=FALSE}
plot_metric(
  ncgh,
  ridge=ridge,
  least_squares=least_squares)
```
